~\Large Increased sexual dimorphism does not evolve in a fossil stickleback following ecological release from fish piscivores

### 2024-10-24 manuscript/paper.tex
finished for the day.
Add Matt's changes to the manuscript.

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{setspace} \setstretch{1.15} \usepackage{float} \floatplacement{figure}{t}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  colorlinks=true,
  linkcolor={cyan},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={cyan},
  pdfcreator={LaTeX via pandoc}}

\title{~\Medium Increased sexual dimorphism does not evolve in a fossil
stickleback following ecological release from fish piscivores}
\author{\large Matthew Stuart\(^{1,2}\), Allison Ozark\(^{3}\), Raheyma
Siddiqui\(^3\), Akhil Ghosh\(^{1,2}\)\\
\large Samantha Swank\(^3\), Michael A. Bell\(^4\), Gregory J.
Matthews\(^{1,2}\), and Yoel E. Stuart\(^{3,+}\)\\
\vspace{-1.1mm}\\
\large \(^1\) Department of Mathematics and Statistics, Loyola
University Chicago, Chicago, IL, USA \vspace{-1.1mm}\\
\large \(^2\) Center for Data Science and Consulting, Loyola University
Chicago, Chicago, IL, USA \vspace{-1.1mm}\\
\large \(^3\) Department of Biology, Loyola University Chicago, Chicago,
IL, USA \vspace{-1.1mm}\\
\large \(^4\) University of California Museum of Paleontology, Berkeley,
CA, USA \vspace{-1.1mm}\\
\large \(^+\) Corresponding:
\href{mailto:ystuart@luc.edu}{\nolinkurl{ystuart@luc.edu}}
\vspace{-1.1mm}}
\date{}

\begin{document}
\maketitle
\begin{abstract}
Theory suggests that populations should expand their habitat- and
resource-use niches when they are freed from interactions with
competitors and predators--so-called ecological release. One way that
this ecological release can manifest is through increased differences
between males and females of the same species as they diverge into
different niches. If divergent niche use drives corresponding divergent
adaptation in in the traits used to exploit the divergent niches, then
theory predicts an increase in sexual dimorphism for those traits,
sometimes called character displacement between the sexes. Using a
dataset of 16 traits collected from 18 temporal samples spread across
16,500 years of evolution of the fossil stickleback fish,
\emph{Gasterosteus doryssus}, we tested the prediction that release from
predation inferred in this system would result in character displacement
between the sexes. We used data from populations of extant stickleback
(\emph{Gasterosteus aculeatus}) to build an imputation model to classify
individual sex in fossil specimens. Then, we quantified the extent of
sexual dimorphism at each temporal sample, and tracked how it changed
through time. \vspace{2mm}\\
\emph{Keywords}: character displacement, character release, evolutionary
time series, \emph{Gasterosteus doryssus}, missing data imputation
\end{abstract}

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2034}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0847}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.7119}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Trait Name
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Trait Code
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Trait Description
\end{minipage} \\
\midrule()
\endhead
Standard Length & \texttt{stl} & Distance from anterior tip of
premaxilla to posterior end of last vertebra (i.e., the llhypural
plate) \\
Dorsal Spine Count & \texttt{mds} & Number of dorsal spines \\
Dorsal Fin Ray Count & \texttt{mdf} & Number of bones in the dorsal fin
posterior to the third dorsal spine (i.e., soft dorsal fin rays) \\
Anal Fin Ray Count & \texttt{maf} & Number of bones in the anal fin
posterior to the anal spine (i.e., soft anal fin rays) \\
Abdominal Vertebra Count & mav & Number of vertebrae anterior to the
first vertebra contacting an anal fin pterygiophore (Aguirre et
al.~2014) \\
Caudal Vertebra Count & \texttt{mcv} & Number of vertebrae posterior to
and including the first vertebra contacting an anal fin pterygiophore
(Aguirre et al.~2014) \\
Pterygiophore Count & \texttt{mpt} & Number of pterygiophores anterior
to but excluding the pterygiophore under the third dorsal spine \\
Pelvic Spine Length & \texttt{lps.sc} & Length from the base of one
pelvic spine at its pelvic girdle articulation to its distal tip \\
Ectocoracoid Length & \texttt{ect.sc} & Length from the anterior to
posterior tips of the shoulder girdle base \\
Pelvic Girdle Length & \texttt{tpg.sc} & Length from the anterior to
posterior tips along midline. If vestigial, the sum of longest axis
along vestiges \\
Cleithrum Length & \texttt{cle.sc} & Length from dorsal tip to ventral
tip on the anterior margin of the shoulder girdle \\
Premaxilla Length & \texttt{pmx.sc} & Length from the anterior tip of
the premaxilla to the distal tip of its ascending process \\
Dorsal Spine Length & \texttt{Ds\#.sc\ \#=1/2/3} & Length from the
dorsal spine base at the pterygiophore to its distal tip \\
Pterygiophore Length & \texttt{lpt.sc} & Distance from anterior to
posterior edges of the pterygiophore immediately before ds3 (when
present) \\
\bottomrule()
\end{longtable}

@@

Briefly, we used multiple imputation (Little and Rubin
(\protect\hyperlink{ref-little2002statistical}{2002})) to impute the sex
of the fossils. Let \(\boldsymbol{W}\) be an
\((n_{extant} + n_{fossil}) \times 1\) vector of the covariate sex of
the stickleback fish and \(\boldsymbol{Y}\) be an
\((n_{extant} + n_{fossil}) \times K\) matrix of the \(K\) phenotypes of
interest. Because the sex of the fossilized stickleback fish is
unobservable, we further define
\(\boldsymbol{W} = (\boldsymbol{W}_{extant}^T,\boldsymbol{W}_{fossil}^T)^T\)
where \(\boldsymbol{W}_{extant}\) and \(\boldsymbol{W}_{fossil}\) are
the \(n_{extant} \times 1\) and \(n_{fossil} \times 1\) vectors of the
observed extant sex and missing fossil sex, respectively. We imputed
missing sex for the fossil data by sampling from the posterior
predictive distribution
\(P(\boldsymbol{W}_{fossil}|\boldsymbol{W}_{extant},\boldsymbol{Y})\)
using the multiple imputation by chained equations (MICE) algorithm
(Buuren and Groothuis-Oudshoorn (\protect\hyperlink{ref-MICE}{2011}))
with predictive mean matching, implemented in R (CITATION).
Traditionally, the choice for the number of completed dats sets is a
relatively small number such as \(M = 5\) or \(M = 10\). However, Zhou
and Reiter (\protect\hyperlink{ref-ZhouReiter2010}{2010}) recommend a
larger number of imputed data sets if users intend on performing
Bayesian analysis after imputation. Therefore, we imputed \(M = 100\)
complete datasets.
\textcolor{red}{PRESUMABLY this also imputes missing data in the other traits? We should probably say that explicitly. -YS}

\textcolor{red}{AKHIL'S METHODS FOR VALIDATION OF THE IMPUTATION MODEL GOES HERE}

@@

The goal of our analysis was to estimate fossil trait means through
time, by sex, while incorporating uncertainty across imputations in the
classification of sex. We used a Bayesian approach to estimate trait
means in part because imputation is the process of sampling from a
posterior predictive distribution of missing data, given observed data.
\textcolor{red}{THIS NEEDS MORE WORK TO JUSTIFY THE BAYESIAN APPROACH, EXPLAINING IT TO MY AUDIENCE. -YS}

\textcolor{red}{MATT, somewhere we need to describe how you handled traits with lots of evolved zeros (ds1, ds2, ds3, lps, tpg) -YS}

@@

Continuous traits will likely have some correlation with standard length
(i.e., allometry) (Huxley
(\protect\hyperlink{ref-Huxley1932}{1932}))(Voje, Bell, and Stuart
(\protect\hyperlink{ref-Vojeetal2022}{2022}))(Voje et al.
(\protect\hyperlink{ref-Vojeetal2014}{2014})). We account for this by
adding an additional parameter, \(\gamma_k\), If \(Y_{k,ij}\) is a
continuous trait, then \begin{align}
{Y}_{k,ij} & \overset{iid}{\sim}\left\{\begin{array}{llll} \mathcal{N}(\mu_{k,ft_i} + \gamma_kY_{K,ij},\sigma_k^2), & W_{ij} = \text{Female} \\ \mathcal{N}(\mu_{k,mt_i} + \gamma_kY_{K,ij},\sigma_k^2), & W_{ij} = \text{Male} \end{array}\right..
\label{eq:cont}
\end{align}

@@

In addition, we modeled two discrete traits, abdominal vertebrae number
(mav) and caudal vertebrae number (mcv) to vary with standard length.
This is because these are vertebral counts that made along the
anterior-posterior access of the fish; these vertebrae contribute to the
length of the fish directly and may evolve in a correlated fashion. If
\(Y_{k,ij}\) is one of the above traits, then \begin{align}
{Y}_{k,ij} & \sim \left\{\begin{array}{ll} GP(\exp\{\mu_{k,ft_i} + \gamma_kY_{K,ij}\},\alpha_k), & W_{ij} = \text{Female} \\ GP(\exp\{\mu_{k,mt_i} + \gamma_kY_{K,ij}\},\alpha_k), & W_{ij} = \text{Male} \end{array}\right..
\label{eq:disc_corr}
\end{align}

For the other discrete counting traits, we set \(\gamma_k = 0\) because
we expect that trait counts are set during early ontogeny and should not
change with size in adult fish. Thus, \begin{align}
{Y}_{k,ij} & \sim \left\{\begin{array}{ll} GP(\exp\{\mu_{k,ft_i}\},\alpha_k), & W_{ij} = \text{Female} \\ GP(\exp\{\mu_{k,mt_i}\},\alpha_k), & W_{ij} = \text{Male} \end{array}\right..
\label{eq:disc_ind}
\end{align} We point out that counting trait means are represented by
\(\exp\{\mu_{k,ft_i}\}\) and \(\exp\{\mu_{k,mt_i}\}\) for ease of
modeling.

In the above model descriptions, \(\mu_{k,ft_i}\) and \(\mu_{k,mt_i}\)
model the time-\(t_i\), sample-specific trait mean \(k\) for females and
males, respectively. To account for the possibility that there is a
time-dependent (i.e., evolutionary trend) trend in \(\mu_{k,ft_i}\) and
\(\mu_{k,mt_i}\), we further set \begin{align}
\mu_{k,gt_i} = \beta_{0,kg} + \beta_{1,kg}t_i + u_{k,gt_i},
\label{eq:mu}
\end{align} for \(g \in \{f,m\}\), where \(\beta_{0,kg}\) and
\(\beta_{1,kg}\) are regression parameters of phenotype \(Y_k\) for each
sex, and \(u_{k,gt_i}\) is the corresponding residual. To account for
potential correlations between the residuals for a given trait \(k\) and
sex \(g\), we fit an Ornstein-Uhlenback (OU) process (Uhlenbeck and
Ornstein (\protect\hyperlink{ref-OUProcess}{1930})). More specifically,
we defined \(du_{k,gt} = u_{k,g(t + dt)} - u_{k,gt}\), the change in
\(u_k,gt\) for a given trait \(k\) and sex \(g\) over a miniscule time
period \(dt\). The OU process is defined as \begin{align}
du_{k,gt} = -\kappa_k u_{k,gt} dt + \tau_k dW_t,
\label{eq:OU_cont}
\end{align} where \(\kappa_k\) is a parameter associated with the
correlation between \(u_{k,gt}\) and \(u_{k,g(t+dt)}\), \(\tau_k\) is
the standard deviation of the OU process, and \(W_t\) is a standard
Brownian motion. As shown in Uhlenbeck and Ornstein
(\protect\hyperlink{ref-OUProcess}{1930}), the closed form solution for
the stochastic differential equation in (\ref{eq:OU_cont}) is
\begin{align}
u_{k,gt_i} \overset{iid}{\sim}\mathcal{N}\left(u_{k,gt_{i-1}}\exp\{-\kappa_k(t_{i} - t_{i-1})\} , \frac{\tau_k^2(1 - \exp\{-2\kappa_k(t_{i} - t_{i-1}))\}}{2\kappa_k}\right)
\label{eq:OU_sol}
\end{align} for \(i = 2,\cdots,T\). In a traditional OU process, the
initial value \(u_{k,gt_1}\) is assumed to be a (potentially unknown)
constant. This unknown constant can therefore be considered a latent
variable, further justifying our Bayesian approach, which better
incorporates prior information and uncertainty from missing data
imputation in the estimate of latent variables.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{setspace} \setstretch{1.15} \usepackage{float} \floatplacement{figure}{t}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  colorlinks=true,
  linkcolor={cyan},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={cyan},
  pdfcreator={LaTeX via pandoc}}

\title{~\Large Increased sexual dimorphism does not evolve in a fossil
stickleback following ecological release from fish piscivores}
\author{\large Matthew Stuart\(^{1,2}\), Allison Ozark\(^{3}\), Raheyma
Siddiqui\(^3\), Akhil Ghosh\(^{1,2}\)\\
\large Samantha Swank\(^3\), Michael A. Bell\(^4\), Gregory J.
Matthews\(^{1,2}\), and Yoel E. Stuart\(^{3,+}\)\\
\vspace{-1.1mm}\\
\large \(^1\) Department of Mathematics and Statistics, Loyola
University Chicago, Chicago, IL, USA \vspace{-1.1mm}\\
\large \(^2\) Center for Data Science and Consulting, Loyola University
Chicago, Chicago, IL, USA \vspace{-1.1mm}\\
\large \(^3\) Department of Biology, Loyola University Chicago, Chicago,
IL, USA \vspace{-1.1mm}\\
\large \(^4\) University of California Museum of Paleontology, Berkeley,
CA, USA \vspace{-1.1mm}\\
\large \(^+\) Corresponding:
\href{mailto:ystuart@luc.edu}{\nolinkurl{ystuart@luc.edu}}
\vspace{-1.1mm}}
\date{}

\begin{document}
\maketitle
\begin{abstract}
Theory suggests that populations should expand their habitat- and
resource-use niches when they are freed from interactions with
competitors and predators--so-called ecological release. One way that
this ecological release can manifest is through increased differences
between males and females of the same species as they diverge into
different niches. If divergent niche use drives corresponding divergent
adaptation in in the traits used to exploit the divergent niches, then
theory predicts an increase in sexual dimorphism for those traits,
sometimes called character displacement between the sexes. Using a
dataset of 16 traits collected from 18 temporal samples spread across
16,500 years of evolution of the fossil stickleback fish,
\emph{Gasterosteus doryssus}, we tested the prediction that release from
predation inferred in this system would result in character displacement
between the sexes. We used data from populations of extant stickleback
(\emph{Gasterosteus aculeatus}) to build an imputation model to classify
individual sex in fossil specimens. Then, we quantified the extent of
sexual dimorphism at each temporal sample, and tracked how it changed
through time. NEED TO FINISH THIS \vspace{2mm}\\
\emph{Keywords}: character displacement, character release, evolutionary
time series, \emph{Gasterosteus doryssus}, missing data imputation
\end{abstract}

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2034}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0847}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.7119}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Trait Name
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Trait Code
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Trait Description
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Standard Length & \texttt{stl} & Distance from anterior tip of
premaxilla to posterior end of last vertebra (i.e., the llhypural
plate) \\
Dorsal Spine Count & \texttt{mds} & Number of dorsal spines \\
Dorsal Fin Ray Count & \texttt{mdf} & Number of bones in the dorsal fin
posterior to the third dorsal spine (i.e., soft dorsal fin rays) \\
Anal Fin Ray Count & \texttt{maf} & Number of bones in the anal fin
posterior to the anal spine (i.e., soft anal fin rays) \\
Abdominal Vertebra Count & \texttt{mav} & Number of vertebrae anterior
to the first vertebra contacting an anal fin pterygiophore (Aguirre et
al.~2014) \\
Caudal Vertebra Count & \texttt{mcv} & Vertebrae including and posterior
to the first that contacts an anal fin pterygiophore (Aguirre et
al.~2014) \\
Pterygiophore Count & \texttt{mpt} & Number of pterygiophores anterior
to but excluding the pterygiophore under the third dorsal spine \\
Pelvic Spine Length & \texttt{lps.sc} & Length from the base of one
pelvic spine at its pelvic girdle articulation to its distal tip \\
Ectocoracoid Length & \texttt{ect.sc} & Length from the anterior to
posterior tips of the shoulder girdle base \\
Pelvic Girdle Length & \texttt{tpg.sc} & Length from the anterior to
posterior tips along midline. If vestigial, the sum of longest axis
along vestiges \\
Cleithrum Length & \texttt{cle.sc} & Length from dorsal tip to ventral
tip on the anterior margin of the shoulder girdle \\
Premaxilla Length & \texttt{pmx.sc} & Length from the anterior tip of
the premaxilla to the distal tip of its ascending process \\
Dorsal Spine Length & \texttt{Ds\#.sc\ \#=1/2/3} & Length from the
dorsal spine base at the pterygiophore to its distal tip \\
Pterygiophore Length & \texttt{lpt.sc} & Distance from anterior to
posterior edges of the pterygiophore immediately before ds3 (when
present) \\
\end{longtable}

@@

Briefly, we used multiple imputation (Little and Rubin
(\protect\hyperlink{ref-little2002statistical}{2002})) to impute the sex
of the fossils. Let \(\boldsymbol{W}\) be an
\((n_{extant} + n_{fossil}) \times 1\) vector of the covariate sex of
the stickleback fish and \(\boldsymbol{Y}\) be an
\((n_{extant} + n_{fossil}) \times K\) matrix of the \(K\) phenotypes of
interest. Because the sex of the fossilized stickleback fish is
unobservable, we further define
\(\boldsymbol{W} = (\boldsymbol{W}_{extant}^T,\boldsymbol{W}_{fossil}^T)^T\)
where \(\boldsymbol{W}_{extant}\) and \(\boldsymbol{W}_{fossil}\) are
the \(n_{extant} \times 1\) and \(n_{fossil} \times 1\) vectors of the
observed extant sex and missing fossil sex, respectively. We imputed
missing sex for the fossil data by sampling from the posterior
predictive distribution
\(P(\boldsymbol{W}_{fossil}|\boldsymbol{W}_{extant},\boldsymbol{Y})\)
using the multiple imputation by chained equations (MICE) algorithm
(Buuren and Groothuis-Oudshoorn (\protect\hyperlink{ref-MICE}{2011}))
with predictive mean matching, implemented in R (CITATION).
Traditionally, the choice for the number of completed dats sets is a
relatively small number such as \(M = 5\) or \(M = 10\). However, Zhou
and Reiter (\protect\hyperlink{ref-ZhouReiter2010}{2010}) recommend a
larger number of imputed data sets if users intend on performing
Bayesian analysis after imputation. Therefore, we imputed \(M = 100\)
complete datasets.
\textcolor{red}{Greg, PRESUMABLY this also imputes missing data in the other traits? We should probably say that explicitly. -YS}
\textcolor{red}{Greg, AKHIL'S METHODS FOR VALIDATION OF THE IMPUTATION MODEL GOES HERE}

@@

The goal of our analysis was to estimate fossil trait means through
time, by sex, while incorporating uncertainty across imputations in the
classification of sex. As we explain in this section, we are
implementing a complex model structure to account for the fact that the
rate of change in the mean of the trait over time may be a non-constant
value. In addition, we are using multiple imputation procedure to
account for the fact that the sex is unobserved in the fossils in
addition to other pieces of information that are missing. We implement a
Bayesian approach because these methods are more effective for analyzing
hierarchical model structures as well as incorporating missing data.

@@

Continuous traits will likely have some correlation with standard length
(i.e., allometry) (Huxley
(\protect\hyperlink{ref-Huxley1932}{1932}))(Voje, Bell, and Stuart
(\protect\hyperlink{ref-Vojeetal2022}{2022}))(Voje et al.
(\protect\hyperlink{ref-Vojeetal2014}{2014})). We account for this by
adding an additional parameter, \(\gamma_k\). In addition, the
continuous traits ds1, ds2, ds3, lps, tpg are unique in that these
traits tend to be lost over time, results in a proportion of zeros in
our empirical dataset. For these random variables, we add binary
variable \(Z_{ij}\), where \(Z_{k,ij} = 1\) means that the trait is
lost, and \(Z_{k,ij} = 0\) means the trait is observed. For all other
continuous traits, we assume \(Z_{k,ij} = 0\)

If \(Y_{k,ij}\) is a continuous trait, then, potentially conditioning on
\(Z_{k,ij} = 0\), \begin{align}
{Y}_{k,ij} & \overset{iid}{\sim}\left\{\begin{array}{llll} \mathcal{N}(\mu_{k,ft_i} + \gamma_kY_{K,ij},\sigma_k^2), & W_{ij} = \text{Female} \\ \mathcal{N}(\mu_{k,mt_i} + \gamma_kY_{K,ij},\sigma_k^2), & W_{ij} = \text{Male} \end{array}\right..
\label{eq:cont}
\end{align} In other words, (\ref{eq:cont}) only considers the subset of
fossils where the given trait is present when performing a data
analysis. For the traits with potential loss, we model the \(Z_{k,ij}\)
via a \(Bernoulli(p_{k,ij})\) distribution where \begin{align}
\text{logit}({p}_{k,ij}) & = \left\{\begin{array}{llll} \alpha_{k,ft_i} + \beta_kY_{K,ij} & W_{ij} = \text{Female} \\ \alpha_{k,mt_i} + \beta_kY_{K,ij}, & W_{ij} = \text{Male} \end{array}\right.,
\label{eq:cont_prob}
\end{align} and logit is a function to convert the probability to the
log-odds, to allow for easier modelling of linear techniques while still
ensuring any probability predictions are guaranteed to be between 0 and
1.

@@

In addition, we modeled two discrete traits, abdominal vertebrae number
(mav) and caudal vertebrae number (mcv) to vary with standard length.
This is because these are vertebral counts that made along the
anterior-posterior access of the fish; these vertebrae contribute to the
length of the fish directly and may evolve in a correlated fashion. If
\(Y_{k,ij}\) is one of the above traits, then \begin{align}
{Y}_{k,ij} & \sim \left\{\begin{array}{ll} GP(\lambda_{k,ft_i} = \exp\{\mu_{k,ft_i} + \gamma_kY_{K,ij}\},\alpha_k), & W_{ij} = \text{Female} \\ GP(\lambda_{k,mt_i} = \exp\{\mu_{k,mt_i} + \gamma_kY_{K,ij}\},\alpha_k), & W_{ij} = \text{Male} \end{array}\right..
\label{eq:disc_corr}
\end{align}

For the other discrete counting traits, we set \(\gamma_k = 0\) because
we expect that trait counts are set during early ontogeny and should not
change with size in adult fish. Thus, \begin{align}
{Y}_{k,ij} & \sim \left\{\begin{array}{ll} GP(\lambda_{k,ft_i} = \exp\{\mu_{k,ft_i}\},\alpha_k), & W_{ij} = \text{Female} \\ GP(\lambda_{k,mt_i} = \exp\{\mu_{k,mt_i}\},\alpha_k), & W_{ij} = \text{Male} \end{array}\right..
\label{eq:disc_ind}
\end{align} We point out that counting trait means are represented by
\(\exp\{\mu_{k,ft_i}\}\) and \(\exp\{\mu_{k,mt_i}\}\). This distinction
helps in our modelling procedure because \(\mu_{k,ft_i}\) and
\(\mu_{k,mt_i}\) are represented on the real number line. This allows
for easier modelling by linear techniques, and that any mean for the
discrete traits is strictly positive. In addition, to ensure proper
sampling from our posterior distribution of interest, we transform
\(\phi_k = \log\left(\frac{\alpha_k - \max_{i,j}(-\lambda_{k,ij}/y_{k,ij})}{1 - \alpha_k}\right)\)
where
\(\lambda_{k,ij} = \left\{\begin{array}{cc} \exp(\mu_{k,ft_i} + \gamma_k Y_{K,ij}) & W_{ij} = \text{Female} \\ \exp(\mu_{k,mt_i} + \gamma_k Y_{K,ij}) & W_{ij} = \text{Male} \end{array}\right.\).
